Bird beak's elaboration and exploration

### 2021-08-23 Analysis/03-elaboration_exploration_analyses.Rmd
Updated first part of the analyses

```{r, eval = TRUE, echo = FALSE}
library(beer)
## Loading the MCMCglmm
load("../Data/processed/combined_results.rda")
## Loading the tree and data
data(demo_data)
data(demo_tree)
## Setting the colour vector
colour_vector <- c("orange", "blue", "darkgreen")
```

@@

```{r, eval = TRUE}
set.seed(42)
## Selecting 1000 random covariance matrices from the MCMCglmm
covar_matrices <- get.covar(combined_results, n = 1000)[1:4]
```

@@

These major axes were calculated as the longest distance within the 95% confidence interval hyper-ellipse from the variance-covariance matrix (i.e. the major axes of the 95% CI ellipse of matrix). <!-- % NC: A small figure here would help -->

@@

```{r}
## Get the centre of each clade
group_centres <- list(
    "gulls"      = colMeans(demo_data[demo_data$clade == "gulls", c(1:4)]),
    "plovers"    = colMeans(demo_data[demo_data$clade == "plovers", c(1:4)]),
    "sandpipers" = colMeans(demo_data[demo_data$clade == "sandpipers", c(1:4)]),
    "phylogeny"  = colMeans(demo_data[, c(1:4)]))

## Calculating the major axes
major_axes <- get.axes(covar_matrices, centre = group_centres)
```

One can visualise these ellipses and major axes in 2D using the `beer::plot.axes` and `beer::plot.ellipses` functions (based on @ellipse):

```{r}
set.seed(51)
## Sub-selecting 5 random VCV matrices for readibility
sub_sample_matrices <- get.covar(combined_results, 5)[1:4]
sub_sample_major_axes <- get.axes(sub_sample_matrices, centre = "intercept")

## Plotting the ellipses
plot.ellipses(sub_sample_matrices, col = c("grey",colour_vector),
              add = FALSE, transparent.scale = 0.5)
plot.axes(sub_sample_major_axes, col = c("grey",colour_vector),
              add = TRUE, transparent.scale = 0.7)
legend("topleft",
       legend = c("phylogeny", levels(demo_data$clade)),
       col = c("grey", colour_vector), lty = 1)
```

We can also visualise all the major axes in relation to the elements in the space.

```{r, fig.height = 10, fig.width = 10}
## Plotting it
plot.space(demo_data,
           col = colour_vector,
           levels = demo_data$clade,
           xlab = "PC1 (90.5%)",
           ylab = "PC2 (6.86%)",
           xlim = c(-2, 2),
           ylim = c(-1, 1))
points(do.call(rbind,group_centres)[,c(1,2)],
       pch = 13, col = c(colour_vector, "grey"))
plot.axes(major_axes["animal"], col = "grey",
          add = TRUE, transparent.scale = 0.01,
          centre = 0)
plot.axes(major_axes["animal:clade1"], col = colour_vector[1],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "gulls", c(1,2)]))
plot.axes(major_axes["animal:clade2"], col = colour_vector[2],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "plovers", c(1,2)]))
plot.axes(major_axes["animal:clade3"], col = colour_vector[3],
          add = TRUE, transparent.scale = 0.01,
          centre = colMeans(demo_data[demo_data$clade == "sandpipers", c(1,2)]))
legend("topleft",
       legend = c("phylogeny", levels(demo_data$clade)),
       col = c("grey", colour_vector),
       lty = c(1,1,1,1),
       pch = c(NA, 19, 19, 19))
```

>>>>

```{r, eval = TRUE, echo = FALSE}
library(beer)
## Loading the MCMCglmm
load("../Data/processed/combined_results.rda")
## Loading the tree and data
data(demo_data)
data(demo_tree)
## Setting the colour vector
colour_vector <- c("orange", "blue", "darkgreen", "grey")
```

@@

```{r, eval = TRUE}
covar_matrices <- MCMCglmm.dispRity(
                  data          = demo_data,
                  posteriors    = combined_results,
                  group         = MCMCglmm.levels(combined_results)[1:4],
                  rename.groups = c("gulls", "plovers", "sandpipers", "phylogeny"))
```

```{r, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5}
## Plotting just the space
sauron.plot(covar_matrices,
            col = colour_vector,
            main = "demo space",
            legend = TRUE)
```

@@

These major axes were calculated as the longest distance within the 95% confidence interval hyper-ellipse from the variance-covariance matrix (i.e. the major axes of the 95% CI ellipse of matrix).

@@

One can visualise these ellipses and major axes in 2D using the `sauron.plot`functions with the `major.axis = mean` and `ellipses = mean` options (based on @ellipse) to display the mean major.axis and the mean ellipses:

```{r, fig.height = 6, fig.width = 6}
sauron.plot(covar_matrices,
            col = colour_vector,
            main = "Average ellipses and\nmajor axes",
            major.axes = mean,
            ellipses = mean,
            legend = TRUE)
```

We can also visualise 100 random major axes and 100 random ellipses:

```{r, fig.height = 12, fig.width = 6}
par(mfrow = c(2,1))
## Plotting 100 axes
sauron.plot(covar_matrices,
            n = 100,
            col = colour_vector,
            main = "100 random major axes",
            major.axes = TRUE,
            legend = TRUE,
            points = FALSE,
            xlim = c(-2, 2),
            ylim = c(-2, 2))
## Plotting 100 ellipses
sauron.plot(covar_matrices,
            n = 100,
            col = colour_vector,
            main = "100 random ellipses",
            ellipses = TRUE,
            legend = TRUE,
            points = FALSE,
            xlim = c(-2, 2),
            ylim = c(-2, 2))
```
